---
title: "STAT 153 Lab 1"
author: "Johnny Hong"
date: "August 29, 2016"
output:
  html_document:
    toc: true
    toc_depth: 3
---

# Introduction
This article is a quick start on using R to deal with time-dependent data.

# Time indices
In R, there are several options for dealing with time indices.

## Date Objects
For dates, the class `Date` in the `base` package can be useful.

### A plotting example
```{r}
start_date <- as.Date("2010-01-01")
end_date <- as.Date("2011-12-31")
dates <- seq(start_date, end_date, by=1)
set.seed(100)
net_worth <- cumsum(rnorm(365 * 2))
plot(dates, net_worth, type="l",
     main="Johnny's Net Worth (Fictional)",
     xlab="Date",
     ylab="Value (in US dollars)")

# Want a more informative x-axis
plot(dates, net_worth, type="l",
     main="Johnny's Net Worth (Fictional)",
     xlab="Date",
     ylab="Value (in US dollars)",
     xaxt="n")
axis.Date(1, at=seq(start_date, end_date, by="3 mon"), format="%Y-%m")
```

### Formatting
Often the date is not presented in the form `YYYY-MM-DD` and directly using
`as.Date()` might result in an error. On the other hand, sometimes we might
want to convert the date to a format other than `YYYY-MM-DD`.
```
as.Date("12/18/1993", format="%m/%d/%Y")
as.Date("September 2, 2013", format="%B %d, %Y")
as.Date("03Aug2014", format="%d%B%Y")
format(Sys.Date(), "%b. %d, %Y")
```
For more information on how to specify the date format, see the documentation for `strptime`.

## POSIXt objects
POSIX stands for "Portable Operating System Interface". There are two subclasses of POSIXt: 1. POSIXct (calendar time): a signed number of seconds since 1970-01-01; 2. POSIXlt (local time): a named list of attributes, including the second (sec), minute (min), hour (hour), day of the month (mday), month (mon), year (year), day of the week (wday), day of the year (yday), and daylight savings time flag (isdst).

```{r}
heart_rate1 <- read.table("http://ecg.mit.edu/time-series/hr.11839")[, 1]
heart_rate2 <- as.vector(read.table("http://ecg.mit.edu/time-series/hr.7257"))[, 1]
numObs <- length(heart_rate1)
# Suppose the first observation was recorded at 4:25pm on Apr 3, 2004. 
times <- as.POSIXct("2004-04-03 16:25:00") + seq(0, len=1800, by=0.5)

summary(heart_rate1)
summary(heart_rate2)
var(heart_rate1)
var(heart_rate2)
plot(heart_rate1 ~ times, type="l",
     xlab="time", ylab="heart rate",
     main="Heart Rates")
lines(heart_rate2 ~ times, col="red")
legend("bottom", legend=c("Subject 1", "Subject 2"), col=c("black", "red"),
       lty=1)
```

## Arithmetics
Doing arithmetics (addition and subtraction) with date-time objects is almost
as easy as doing arithmerics with ordinary numbers. The package `lubridate`
facilitates some of the operations.

```{r}
# POSIXct objects
ex1 <- as.POSIXct("2014-02-03 16:00:00")
# Add 10 seconds (note the by default 10 means 10 seconds)
ex2 <- ex1 + 10
print(ex2)
# Add 10 hours
ex3 <- ex1 + 10 * 60 * 60
print(ex3)
# Add 10 days
ex4 <- ex1 + 10 * 24 * 60 * 60
print(ex4)

library(lubridate)
# Add 10 hours
ex5 <- ex1 + hours(10)
print(ex5)
ex6 <- ex1 + days(10)
print(ex6)
```

# `ts` objects
The `ts` object is useful for analyzing time series with regular time spacings.
```{r}
print(nottem)
class(nottem)
tsp(nottem) # start time, end time, frequency
plot(nottem, main="Average air temperature at Nottingham Castle",
  xlab="Time",
  ylab="Temperature (Fahrenheit)")
# We will discuss this in the future.
plot(stl(nottem, "per"))
```

# `ggplot2`
For visualization, besides the `base` graphics, the R package `ggplot2`,
developed by Hadley Wickham, are often used. While `ggplot2` is a bit harder
to learn compared to `base` graphics, the resulting graphics are often more
visually appealing. For a slightly more in-depth introduction to `ggplot2`, see [here](http://jcyhong.github.io/ggplot_demo.html).

```{r}
library(ggplot2)
library(zoo)

# Redo heart rate example.
heart_rate_data <- data.frame(times, heart_rate1, heart_rate2)
colnames(heart_rate_data) <- c("Time", "Subject 1", "Subject 2")
library(reshape2)
heart_rate_data_long <- melt(heart_rate_data,
                             id.vars="Time")
ggplot(heart_rate_data_long,
        aes(x=Time, y=value, color=variable)) +
  geom_line() +
  xlab("time") +
  ylab("heart rate") +
  scale_color_discrete(guide=guide_legend(title = "Subject")) +
  ggtitle("Heart rate time series") +
  theme(
    plot.title = element_text(size=24),
    axis.title.x = element_text(size=14),
    axis.title.y = element_text(size=14),
    axis.text.x = element_text(size=12),
    axis.text.y = element_text(size=12),
    legend.title = element_text(size=14),
    legend.text = element_text(size=12)
  )

economics_mod <- economics
economics_mod$unemploy_ma <- rollmean(economics$unemploy, k=5,
                              fill=list(NA, NULL, NA))
ggplot(economics_mod, aes(x=date)) +
  geom_line(aes(y=unemploy, color="reality")) +
  geom_line(aes(y=unemploy_ma, color="moving average")) +
  scale_colour_manual("Lines", values=c("reality"="black",
                                        "moving average"="red")) +
  ggtitle("US Unemployment") +
  xlab("Unemployment (in thousand)") + 
  ylab("Date")
```

# Other useful packages
There are many more ways in R to deal with time series data. For irregularly spaced
time series, consider the packages `zoo`, `xts`, and `timeSeries`.

# Reference
1. [ggplot2: Elegant Graphics for Data Analysis](http://www.springer.com/us/book/9780387981406) by Hadley Wickham
2. [Working with Financial Time Series Data in R](https://faculty.washington.edu/ezivot/econ424/Working%20with%20Time%20Series%20Data%20in%20R.pdf) by Eric Zivot (2014).
